---
title: "A brief introduction to web scraping using `rvest`"
date: "`r format(Sys.Date(), '%m-%d-%Y')`"
output: html_document
---

```{r opts, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Outline

- *What's our goal?*
    - Basic introduction to web scraping
    - Just enough of the R language to grasp the basics
    - Be able to follow along with a program designed to get data from the internet

- *What is web scraping?*
    - good practice, robots.txt
    - HTML elements, XPaths, looking at a website's 'source'

- *Web scraping in R*
    - quick overview of the R environment, data types
    - primary functions used with the `rvest` package
    
- *Applied example: gathering job postings from labor.gov*

## What is web scraping?

- Going beyond copy/paste

- Using software to systematically gather data from websites
    - Copying information posted online for later review/analysis
    - Can be done manually
    - Or through an automated process ("crawlers", "bots", "spiders")

- Lots of terms & names for describing the same thing:
    - "web scraping" or "scraping"
    - "web harvesting"

## Good practice & responsible scraping

- Read each website's ToS and policies
    - May not be permissible for you to scrape certain data
        - copyright laws
        - trespassing
        - violation of the computer fraud and abuse act (CFAA)
    - Can get your IP Address banned
        - make you sad
        - make your institution mad
    - Public data generally okay to gather, but be conscious
    - Reaching out to web administrators is often a good idea
- Read each website's robots.txt
    - Example: https://en.wikipedia.org/robots.txt
    - How to understand a robots.txt file: https://en.wikipedia.org/wiki/Robots_exclusion_standard

## Using R: packages & workspace

- 'packages' are bundles of code that other people have written for convenient use later
- they contain functions and data that are meant to be used across different projects
- start by running the following command in your console:
    - `install.packages("tidyverse")`

```{r init}
# install packages once, load packages every session
# installing *is not the same* as loading 
# load packages by calling the `library()` function on a package name
library(tidyverse)
```

## Using R: data types (1/)

Quick look at how data is stored and represented in R (you can use the `class()` function on an object to find out what it is)

- *vectors* can be thought of single columns of a spreadsheet
    - most basic type of data
    - have one dimension: length (how many items in the vector)
    - only one type of data contained in them
        - numeric (numbers, either integer values or decimals)
        - character (text based data)
        - logical (`TRUE` or `FALSE`)
        - missing values (represented by `NA` in R)
        - factor (won't cover in detail, just be aware they exist)
    - if you try to mix different types, R will try to convert ("coerce") them into a compatible type

- *lists*
    - a little more advanced
    - used to store sets of vectors or other forms of data (even lists of lists!)
    - as many items in a list as you want
    - each item in the list can be a different type of data
    - and have different dimensions (rows, lengths, etc.)

- *data frames*
    - technically a list, but probably feels more familiar
    - rows and columns (2 dimensions)
    - we <3 data frames

## Using R: data types (2/)

```{r use-r-2}
# vectors
letters[1:12]

# lists
list(my_letters = letters[1:5], my_numbers = 1:10)

# data frames
head(mpg)
```

## Using R: operators (3/)

Operators. One note: `#` marks lines as comments

```{r use-r-3}
# creating sequences using the colon
1:5

5:1

# doing math
3 * 2

1 + 1

3 - 2

4 / 8

4 ^ 2

1:10 - 1 # being fancy

# logical tests
2 == 2

2 != 4

TRUE & TRUE

TRUE | FALSE

# storing data as variables for later use
# <- is the *assignment operator*
x <- 3
y <- 22

(x * y) + 3

# getting data out of an object
mpg_cty <- mpg$cty
```

## Using R: functions (4/)

- Just like in high-school math, functions have a name, and take arguments
- Arguments go inside parentheses

```{r use-r-4, warning = TRUE}
mean(mpg_cty) # find an average of a vector of data
mean(letters) # can't find the average of the alphabet...

table(mpg$class) # the $ operator can be used to access columns in data frames

# ? can be used to pull up help pages about functions, or built-in data
?mean
```

## The `rvest` package

We're going to look at the wikipedia entry for the [gini coefficient](https://en.wikipedia.org/wiki/Gini_coefficient) as an example. We'll use the 'inspect' feature from the chrome browser to look at the structure of the page to get an XPath that can be targeted by functions from `rvest`.

```{r rvest-1}
# load the rvest package that came installed with the `tidyverse` bundle
library(rvest)

# grab the html from the web
gini <- read_html("https://en.wikipedia.org/wiki/Gini_coefficient")

# after we've looked at the page's source, we can get the xpath we want
tabl <- html_nodes(gini, xpath = '//*[@id="mw-content-text"]/div/table[1]')

html_text(tabl)  # a little more work might be needed
html_table(tabl) # but, this is actually a table, so we can take a shortcut
```

## More to `rvest`, but these are the main pieces

You can do a lot more to automate processes with `rvest` (like submitting data into forms, or pulling out more complicated information from HTML files), but we won't need to go into too much more detail.

I've gotten most projects done with the following:

- `read_html()`
- `html_node()` or `html_nodes()`
- `html_text()` or `html_table()`

## An applied example: getting H-2A postings from the DoL

```{r}
```
